---
title: "Introduction to rangejoin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{intro}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction to rangejoin
This package provides a fast and efficient way to perform range joins in R. A range join is a type of join operation where you match rows from two datasets based on whether the values in one dataset fall within a specified range of values in another dataset.
The `rangejoin` package is designed to handle large datasets efficiently, leveraging data.table for fast operations. It supports both left and inner joins, allowing you to combine datasets based on range conditions.
# Installation
You can install the `rangejoin` package from CRAN using the following command:

```{r setup}
library(rangejoin)
```

# Methodological Logic
The `rangejoin` package is designed to efficiently join two datasets based on range conditions, rather than exact matches. This is useful in scenarios such as time intervals, genomic ranges, or any case where you want to match records that overlap or fall within a certain range.

The `split_data` function is used to partition a dataset into non-overlapping intervals or groups, which can be helpful for preprocessing before performing a range join. The `range_join` function then performs the actual join, matching rows from one dataset to another where a value (or range) from one falls within the range of the other.

# Demo Data
The package provides a `demo_data` dataset for illustration. Let's load and inspect it:

```{r demo-data}
library(rangejoin)
data(demo_data)
head(demo_data)
```

# Using `split_data`
Suppose we want to split the demo data into groups based on a column (e.g., `group`). The `split_data` function helps with this:

```{r split-data}
split_list <- split_data(demo_data, by = "group")
str(split_list)
```

This returns a list of data.tables, each corresponding to a unique value in the `group` column.

# Using `range_join`
Now, let's demonstrate a range join. Suppose we have two datasets: `x` and `y`, each with `start` and `end` columns representing intervals. We'll join them where the intervals overlap.

```{r range-join}
# Create two example datasets from demo_data
x <- demo_data[group == 1, .(id, start, end)]
y <- demo_data[group == 2, .(id, start, end)]

# Perform a range join (inner join by default)
result <- range_join(x, y, by = c("start", "end"))
head(result)
```

This will return rows where the intervals in `x` and `y` overlap. You can also specify the type of join (e.g., `type = "left"`).

# Summary
The `rangejoin` package provides efficient tools for working with range-based data joins, making it easy to handle complex interval matching tasks in R.
