---
title: "Introduction to rangejoin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rangejoin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = ""
)
```

This package provides a fast and efficient way to perform range joins in R. A range join is a type of join operation where you match rows from two datasets (for example, one containing treatment cases and another containing control cases) in such a way that matched values fall within a specified range.

# Exported functions

The `split_data()` function is used to partition a dataset into non-overlapping intervals or groups, which can be helpful for preprocessing before performing a range join. The `range_join()` function then performs the actual join, matching rows from one dataset to another where a value (or range) from one falls within the range of the other. Finally, the `find_matches()` function helps visually inspect the matches found by `range_join()`.

# Example

The package provides a `demo_data` dataset for illustration. Let's load and inspect it:

```{r demo-data}
library(rangejoin)
data(demo_data)
head(demo_data)
```

Suppose we want to split the demo data randomly into case and control groups. The `split_data` function helps with this:

```{r split-data}
set.seed(3) # necessary since the splitting is random by default
(split_list <- split_data(demo_data))
```

Obs.: `demo_data` already contains a `case_control` column, which can be used to deterministically split the data, but here we demonstrate the random splitting.

Next, we can perform a range join using the `range_join()` function. Assuming we want to assign controls to each case on `split_list` in such a way that the age of the controls are within [-3, 3] years of the cases:

```{r range-join}
range_joined <- range_join(
  split_list, max_ctrl_per_case = 2L, join_var_range = c(-3, 3)
)
```

Notice the necessity to specify `max_ctrl_per_case` to limit the number of controls assigned to each case.

Finally, we can visualize the matches using the `find_matches()` function. To keep the output short, we will select a few cases to inspect:
```{r find-matches}
find_matches(range_joined, case_list = c(2, 90, 100, 16))
```

# Summary methods

The package also provides summary methods for the results of the range join. You can use `summary()` to get an overview of the matches found:

```{r summary-method}
summary(split_list)
summary(range_joined)
```
